---
title: "RLab2"
author: "pearl_kuuridong_SR2"
date: "2025-11-28"
output: html_document
---

```{r setup, include=FALSE}
library(ggplot2)
library(lme4)
library(lmerTest)
library(merTools)
data <- read.csv('growth_data.txt', sep='\t')
data$child <- as.factor(data$child)
data$treatment <- as.factor(data$treatment)
```

```{r}

selected_children <- c(90211, 90070, 80369)

# Filter manually using base R
three_children <- data[data$child %in% selected_children, ]
```

```{r}
data$height[is.na(data$height)] <- median(data$height, na.rm=T)
data$age[is.na(data$age)] <- median(data$age, na.rm=T)
non_missing <- data$treatment[!is.na(data$treatment)]

class_probs <- prop.table(table(non_missing))

n_na <- sum(is.na(data$treatment))

set.seed(123)
random_impute <- sample(
  names(class_probs), 
  size = n_na, 
  replace = TRUE, 
  prob = class_probs
)

data$treatment[is.na(data$treatment)] <- random_impute
```

```{r}
names(data)

```


```{r}
# Plot using ggplot2
ggplot(three_children, aes(x = age, y = height, colour = child)) +
  geom_line(size=1) +
  geom_point() +
  theme_minimal() +
  labs(title="Growth Curves for Selected Children",
       x="Age", y="Height")
```
# COMMENT:
# The plot shows individual growth curves. Children generally increase in height with age
# but the slopes differ, indicating individual variability. Some children may respond 
# differently depending on treatment or biological differences.
# ============================
# (b) Full data plotted in three panels by treatment
# ============================
```{r}
ggplot(data, aes(x = age, y = height, group = child)) +
  geom_line(alpha=0.4) +
  facet_wrap(~ treatment) +
  theme_minimal() +
  labs(title="All Growth Curves by Treatment Group",
       x="Age", y="Height")
```
# COMMENT:
# Each treatment panel shows variability in growth trajectories. 
# Some treatments exhibit steeper growth slopes or higher variability.
# This suggests treatment may influence growth rate.


# ============================
# (c) Linear mixed model: random intercept
# ============================

```{r}
model_intercept <- lmer(height ~ age + treatment + (1 | child), data=data)
summary(model_intercept)
```

# COMMENT:
# Random intercept captures child-specific baseline height.
# Fixed effects estimate overall effect of age and differences between treatments.
# Treatment may shift average height but slopes remain common across children.


# ============================
# (d) Linear mixed model: random slope
# ============================

```{r}
model_slope <- lmer(height ~ age + treatment + (0 + age | child), data=data)
summary(model_slope)

```

# COMMENT:
# Random slope allows each child to have their own growth rate.
# Age slope varies significantly between children if random slope variance > 0.
# No random intercept in this model.


# ============================
# (e) Random intercept + random slope
# ============================

```{r}
model_full <- lmer(height ~ age + treatment + (age | child), data=data)
summary(model_full)

# Extract random effects with confidence intervals
re <- REsim(model_full)

# Plot random effects
plotREsim(re)

```


# COMMENT:
# This model captures both baseline height differences (intercept)
# and differences in growth rates (slope) across children.
# Confidence intervals show how much each child's random intercept and slope deviate
# from the population average. Wider intervals indicate greater uncertainty.


# ============================
# (f) Compare model in (e) vs model in (c)
# ============================

```{r}
anova(model_intercept, model_full)
```

# COMMENT:
# Likelihood ratio test:
# If p-value < 0.05, the full model (random intercept + slope) fits significantly better.
# This means children differ not only in baseline height but also in growth rate.
# If p-value > 0.05, the simpler random intercept model is adequate.

```{r}

```

