---
title: "pearl_kuuridong_SR1"
author: "pearl_kuuridong_SR1"
date: "2025-11-14"
output: pdf_document
---

```{r}
data <- mtcars
attach(data)

head(data)
```

```{r}
columns <- colnames(data)
columns
```

# Question One

We have the correlations below:
```{r}
for (i in 1:(length(columns)))
{
  for (j in 1:length(columns))
  {
    if (i != j)
    {
      col1 <- columns[i]
      col2 <- columns[j]
      correlation <- cor(data[[col1]], data[[col2]])
      print(paste('The correlation between', col1, 'and', col2, 'is', correlation))
    }
  }
}
```

We have the plots below:
```{r}
for (i in 1:(length(columns)))
{
  for (j in 1:length(columns))
  {
    if (i != j)
    {
      col1 <- columns[i]
      col2 <- columns[j]
      plot(data[[col1]], data[[col2]], xlab=col1, ylab=col2, main = paste('Plot of',col2, 'against', col1))
    }
  }
}
```
In particular, we plot only the graphs for only mpg against all the other variables

```{r}
for (i in 2:length(columns))
{
  col <- columns[i]
  plot(data[['mpg']], data[[col]], xlab='mpg', ylab=col, main=paste('Plot of', col, 'against mpg'), sub = paste('Correlation = ', cor(data[['mpg']], data[[col]])))
  plot(data[[col]], data[['mpg']], xlab=col, ylab='mpg', main=paste('Plot of mpg against', col), sub = paste('Correlation = ', cor(data[[col]], data[['mpg']])))
}
```
and in particular we look at only the correlations between mpg and the other variables
```{r}
for (i in 2:length(columns))
{
  col <- columns[i]
  correlation <- cor(data[['mpg']], data[[col]])
  print(paste('The correlation between', 'mpg', 'and', col, 'is', correlation))
  new_correlation <- cor(data[[col]], data[['mpg']])
  print(paste('The correlation between', col, 'and mpg is', new_correlation))
}
```

# Question Two
```{r}
disp_model = lm(mpg ~ disp, data=data)
drat_model = lm(mpg ~ drat, data=data)
wt_model = lm(mpg ~ wt, data=data)
```

```{r}
summary(disp_model)
confint(disp_model, level = 0.95)
```

```{r}
summary(drat_model)
confint(drat_model, level = 0.95)
```

```{r}
summary(wt_model)
confint(wt_model, level = 0.95)
```
#Question Three
```{r}
plot(disp_model$residuals, ylab='residual_values', main='Plot of Residuals of Fitted Values With disp As Covariate')
abline(h = 0, lty = 2, lwd = 0.9)
```
```{r}
plot(drat_model$residuals, ylab='residual_values', main='Plot of Residuals of Fitted Values With drat As Covariate')
abline(h = 0, lty = 2, lwd = 0.9)
```
```{r}
plot(wt_model$residuals, ylab='residual_values', main='Plot of Residuals of Fitted Values With wt As Covariate')
abline(h = 0, lty = 2, lwd = 0.9)
```
# Question Four
```{r}
predicted_mpg_disp <- predict(disp_model, newdata = data.frame(disp=290))
predicted_mpg_disp
```

```{r}
predicted_mpg_drat <- predict(drat_model, newdata = data.frame(drat=3.25))
predicted_mpg_drat
```

```{r}
predicted_mpg_wt <- predict(wt_model, newdata = data.frame(wt=2.950))
predicted_mpg_wt
```

# Question Five

We define functions to compute the estimators given the data 
```{r}
beta_one_hat <- function(y, x){
  x_bar <- mean(x, na.rm=T)
  sxx <- sum((x - x_bar)^2)
  sxy <- sum(y*(x-x_bar))
  
  return(sxy/sxx)
}

beta_zero_hat <- function(y, x, beta_one){
  y_bar <- mean(y, na.rm=T)
  x_bar <- mean(x, na.rm=T)
  
  beta_zero <- y_bar - (beta_one*x_bar)
  return(beta_zero)
}


mse <- function(y, x, beta_one, beta_zero){
  ss_res <- sum((y - beta_zero - (beta_one*x))^2, na.rm=T)
  n <- length(y)
  return(ss_res/(n-2))
}
```

We compute the estimates for each regression model
```{r}
disp_beta_one <- beta_one_hat(mpg, disp)
disp_beta_zero <- beta_zero_hat(mpg, disp, disp_beta_one)
disp_mse <- mse(mpg, disp, disp_beta_one, disp_beta_zero)

disp_beta_one; 
disp_beta_zero;
disp_mse
```
```{r}
drat_beta_one <- beta_one_hat(mpg, drat)
drat_beta_zero <- beta_zero_hat(mpg, drat, drat_beta_one)
drat_mse <- mse(mpg, drat, drat_beta_one, drat_beta_zero)

drat_beta_one; 
drat_beta_zero;
drat_mse
```

```{r}
wt_beta_one <- beta_one_hat(mpg, wt)
wt_beta_zero <- beta_zero_hat(mpg, wt, wt_beta_one)
wt_mse <- mse(mpg, wt, wt_beta_one, wt_beta_zero)

wt_beta_one; 
wt_beta_zero;
wt_mse
```

# Question Six
```{r}
confint(disp_model, level = 0.95)
```

```{r}
confint(drat_model, level = 0.95)
```

```{r}
confint(wt_model, level = 0.95)
```

# Question Seven
```{r}
plot(fitted(disp_model), residuals(disp_model), main = "Residuals vs Fitted of mpg ~ disp")
abline(h = 0, lty=2, lwd=0.9)
```
```{r}
plot(fitted(drat_model), residuals(drat_model), main = "Residuals vs Fitted of mpg ~ drat")
abline(h = 0, lty=2, lwd=0.9)
```

```{r}
plot(fitted(wt_model), residuals(wt_model), main = "Residuals vs Fitted of mpg ~ wt")
abline(h = 0, lty=2, lwd=0.9)
```

