---
title: "Introduction to R Programming"
author: "Pearl Kuuridong"
date: '`r Sys.Date()`'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Operators, variables and data types in R
To insert a new chunk, use the shortcut CTRL + Alt + I
```{r} 
rm(list=ls())
# This block is called a chunk of code. It works like the usual markdown format
source("R_day1.R")
```
```{r}
set.seed(123) #reproducibility (fixing the randomness)
dice = sample(6, size=100, replace=TRUE)
dice

length(dice)

# filter all the 6s out of dice use the logical operator ==.
# In that regard R vectors are exactly numpy arrays
dice[dice==6]

# use the sum function to calculate the number of 6s in the variable dice
sum(dice==6)
# you can also do something like this:
length(dice[dice==6])

# Use table function to get counts for all categories
table(dice)

# Relative frequency:
# table(dice)/length(dice)
# The above is the same as
prop.table(table(dice))
```

### Generate a vector using `rep`
```{r}
# Repeat some value `10`, 100 times.
# x <- rep(10,100)
(names3x <- rep("Pearl", 3))
```


### This is a subsection
**Exercise:**
<!-- Write an R code to generate 1 1 1 1 0 0 0 0 0 1 2 3 4 6 5 4 3 2 0 0 0 0 -->
```{r}
c(rep(c(1,0), c(4,5)), 1:4, 6:2, rep(0,4))
```

### Operations on vectors


```{r}
vect1 <- c(2,3,5,6,8)
vect2 <- c(1,3,5,8,10)

vect1 + vect2 # add
vect1 * vect2 #element-wise multiplication
vect1 / vect2 #element-wise division
vect1 %*% vect2 #dot product

k= 2
k* vect1

vect1 ** k
vect1 %% 2 #element-wise division
vect1 - k
vect1 + k
vect1 %/% k

mean(vect1)
min(vect1)
max(vect1)
range(vect1)
summary(vect1)[]
```

### Vectors from distributions
#### Normal distribution
A variable $X$ is normally distributed or $X\sim\mathcal{N}(\mu, \sigma^2)$, if: 
$$f_X(x) = \frac{1}{\sqrt(2\pi\sigma^2)}\exp\left\{\frac{-(x-\mu)^2}{2\sigma^2}\right\}, \quad x\in\mathbb{R}$$
```{r}
set.seed(123)
x_norm <- rnorm(150, 60, 10)
hist(x_norm,)
abline(v=mean(x_norm), col="red", lwd=0.25, lty=2) 
```

### Poisson distribution
```{r}
?rpois
x_pois <- rpois(n=150, 2.5)
hist(x_pois,)
abline(v=mean(x_pois), col="red", lwd=0.25, lty=2) 
```


```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
